---
title: "ClassSurvey"
author: "Harry"
date: "2025-08-27"
output: html_document
---

```{r}
# install.packages("readr") 
library(readr)
```

1. Import class one survey data from our Github site (https://github.com/kijohnson/Advanced-Data-Analysis) (The dataset is called ‘Class 1 Survey Fall 2025 2.csv’), calling the R dataframe that you create C1survey.


```{r}
ClassSurv <- read_csv("https://github.com/kijohnson/Advanced-Data-Analysis/raw/refs/heads/main/Class%201%20Survey%20Fall%202025%202.csv")
```
2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.
```{r}
dim(ClassSurv)
```
3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed. Hints: You can use either the rename_with function or the names function. The list of names for time is:
```{r}
install.packages("dplyr")
library(dplyr)
```
```{r}
colnames(ClassSurv)
```


```{r}
names(ClassSurv) <- c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```

4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.
```{r}
num_factors <- sum(sapply(ClassSurv, is.factor))
num_integers <- sum(sapply(ClassSurv, is.integer))
num_numerics <- sum(sapply(ClassSurv, is.numeric))
num_characters <- sum(sapply(ClassSurv, is.character))
```

```{r}
type_counts <- c(
  Factors    = sum(sapply(ClassSurv, is.factor)),
  Integers   = sum(sapply(ClassSurv, is.integer)),
  Numerics   = sum(sapply(ClassSurv, is.numeric)),
  Characters = sum(sapply(ClassSurv, is.character))
)

# Print counts directly
print(type_counts)
```

a. Using code, check the bday and bmonth variables for any unusual or missing values (hint: use table function). If you find any,
```{r}
table(ClassSurv$bday, useNA = "ifany")
table(ClassSurv$bmonth, useNA = "ifany")

```
 b. describe what you will do with the unusual values in a comment before or immediately following the code that does it,
If unusual values are found, we will replace them with NA ,so they can be treated as missing rather than incorrect data.
```{r}
unique(ClassSurv$bday[!(ClassSurv$bday %in% 1:31)])
unique(ClassSurv$bmonth[!(ClassSurv$bmonth %in% 1:12)])
ClassSurv$bday[!(ClassSurv$bday %in% 1:31)] <- NA
ClassSurv$bmonth[!(ClassSurv$bmonth %in% 1:12)] <- NA
```

c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
ClassSurv$bday <- as.numeric(ClassSurv$bday)
ClassSurv$bmonth <- as.numeric(ClassSurv$bmonth)

```

```{r}
median_bday <- median(ClassSurv$bday, na.rm = TRUE)
median_bmonth <- median(ClassSurv$bmonth, na.rm = TRUE)
median_bday
median_bmonth
```

6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30).
```{r}
ClassSurv$bseason <- with(ClassSurv,
  ifelse(bmonth %in% c(12, 1, 2), "Winter",
  ifelse(bmonth %in% 3:5, "Spring",
  ifelse(bmonth %in% 6:8, "Summer",
  ifelse(bmonth %in% 9:11, "Fall", NA)))))
```
```{r}
ClassSurv$bseason <- factor(ClassSurv$bseason, 
                            levels = c("Winter", "Spring", "Summer", "Fall"))
```

b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct 

```{r}
table(ClassSurv$bmonth, ClassSurv$bseason)
```
```{r}
ClassSurv$bmonth <- factor(ClassSurv$bmonth, levels = 1:12)
table(ClassSurv$bmonth, ClassSurv$bseason)
```

c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.
```{r}
tab <- table(ClassSurv$bmonth, ClassSurv$bseason)
addmargins(tab)
```
Winter = 8
Spring = 15
Summer = 13
Fall = 12


7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).

Which birth season has the highest number of medical doctors?
Variables - bseason. highest_educ_levelt

a, I created a cross table of brith season agaiinst highest_educ_level
```{r}
season_vs_education <- table(ClassSurv$bseason, ClassSurv$highest_educ_level, useNA = "ifany")
```

b. I used the code below to produce the table to see raw counts
```{r}
season_vs_education
```
Answer = Fall 



